---
title: "PercAct_Exam"
author: "Laurits Lyngbæk"
date: "2022-12-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Cleaning data
### Read packages

```{r - load packages}
pacman::p_load(tidyverse,mousetrap,readbulk)
```


### Load packages as mousetrap_object
```{r}
df <- read_opensesame(directory = "./Data/", extension = ".csv") #Load all files in folder with .csv ending
```

```{r - non-mousetracking dataframe}
# checking percent correct:
df %>%
    group_by(subject_nr) %>% 
    summarise(round(sum(correct)/22*100))
```

```{r check incorrect}
# What trial type is the errors from
df %>% 
    filter(correct == 0) %>% 
    select(Trial_Type, count_logger)


df %>% 
    filter(correct == 0) %>% 
    group_by(Trial_Type) %>% 
    summarize(n())


#What sizes are most often mistaken.
df %>% 
    filter(correct == 0) %>% 
    mutate(size_diff = abs(Size_1-Size_2)) %>% 
    group_by(size_diff) %>% 
    summarize(n())

```


```{r}
#Converting only the correct raw mouse tracking data into mousetrap data
df <-subset(df, df$correct == 1)
df <-subset(df, df$Data_Type_Sort == 1)
mt<- mt_import_mousetrap(df) #tranform df to mousetrap object
rm(df) #clean variables
```


```{r}
#excluding the initation times for the trials
mt <- mt_exclude_initiation(mt)
#mirror-symmetric mapping of movements
mt <- mt_remap_symmetric(mt, use = "trajectories", remap_xpos = "left")
#Aligning the starting and end position of all the trials (Might need to change this one -
 #it space-normalizes the data.)
mt <- mt_align_start_end(mt)
```

```{r}
#standardizing the time variable for the trials into chunks in order for the data to be comparable between trials
mt <- mt_time_normalize(mt)
#calculating new coloun for the metrics (MAD, AD) used below
mt <- mt_measures(mt)
```

The metrics are as follows:
  -> MAD describes the maximum absolute deviation from the direct path connecting the start and end point of the trajectory 
    (if going in a straight line). Meaning that a value of 0 would be ideal.
    
  -> AD denotes the average deviation from the the direct path.
  
  -> AUC denotes the area under the curve; meaning the geometric area between the actual trajectory and the direct path where the       area below the the direct path has been subtracted
  
  -> xpos_flips denotes the average number of directional changes along the x-axis
  
  -> RT denotes the average response time for the trials

```{r}
# The function is used to aggregate mouse movement data from multiple trials or subjects into a single dataframe.
mt_aggregate(
  mt, use = "measures",
  use_variables = c("MAD", "AD", "AUC", "xpos_flips", "RT"),
  use2_variables = "Trial_Type",
  subject_id = "subject_nr"
)


```
```{r}
#plotting all trial curves with colors based on the condition
mt_plot(
  mt, use = "tn_trajectories",
  x = "xpos", y = "ypos", color = "Trial_Type") +
  theme_minimal() + labs(title = "Aligned time-normalized mouse-tracking data", 
                           x = "Position (x)",
                           y = "Position (y)")
```

```{r}
  #plotting the aggregated curves for the two conditions
mt_plot_aggregate(
  mt, use = "tn_trajectories",
  x = "xpos", y = "ypos", color = "Trial_Type",
  subject_id = "subject_nr") +
  theme_minimal() + labs(title = "Aligned aggregated time-normalized mouse trajectories", 
                           x = "Position (x)",
                           y = "Position (y)")
```
```{r}



```

