---
title: "PercAct_Exam"
author: "Laurits Lyngbæk"
date: "2022-12-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Cleaning data
### Read packages

```{r - load packages}
pacman::p_load(tidyverse,mousetrap,readbulk,lme4, stats)

```


### Load packages as mousetrap_object
```{r}
df <- read_opensesame(directory = "./Data/", extension = ".csv") #Load all files in folder with .csv ending
```

```{r - non-mousetracking dataframe}
#Creating new condition names

df <- df %>%  
    mutate(Trial_Type = ifelse(Trial_Type == "Incongurent", "Incongruent", Trial_Type)) %>% 
    mutate(Trial_Type = ifelse(Trial_Type == "Congurent", "Exaggerated", Trial_Type)) %>% 
    mutate(Trial_Type = ifelse(Trial_Type == "Control", "Congruent", Trial_Type))

df_wrong <- df %>% 
    filter(correct == 0)



# checking percent correct:
df %>% 
    group_by(subject_nr) %>% 
    count(gender) %>% 
    ungroup() %>% 
    count(gender)


df$Trial_Type
df %>% 
    select(gender, age) 


df_mean <- tapply(df$age, df$gender, mean)
df_sd <- tapply(df$age, df$gender, sd)

df_mean
df_sd
```


```{r check incorrect}
df %>% 
    filter(correct == 0) %>%
    # filter(Data_Type != "Training") %>% 
    select(Trial_Type, count_logger) %>% 
    group_by(Trial_Type) %>% 
    summarise(n())


1-(26+10+15)/594
```


```{r check incorrect}
df %>% 
    filter(correct == 0) %>% 
    mutate(size_diff = abs(Size_1-Size_2)) %>% 
    group_by(size_diff) %>% 
    summarize(n())
```


```{r}
#Converting only the correct raw mouse tracking data into mousetrap data
incorrect_mt <-  mt_import_mousetrap(df)

df <-subset(df, df$correct == 1)
df <-subset(df, df$Data_Type != "Training")
mt<- mt_import_mousetrap(df) #tranform df to mousetrap object

```


```{r}
#excluding the initation times for the trials
mt <- mt_exclude_initiation(mt)
#mirror-symmetric mapping of movements
mt <- mt_remap_symmetric(mt, use = "trajectories", remap_xpos = "left")
#Aligning the starting and end position of all the trials (Might need to change this one -
 #it space-normalizes the data.)
mt <- mt_align_start_end(mt)
```

```{r}
#standardizing the time variable for the trials into chunks in order for the data to be comparable between trials
mt <- mt_time_normalize(mt)
#calculating new coloumn for the metrics (MAD, AD) used below
mt <- mt_measures(mt)
```


The metrics are as follows:
  -> MAD describes the maximum absolute deviation from the direct path connecting the start and end point of the trajectory 
    (if going in a straight line). Meaning that a value of 0 would be ideal.
    
  -> AD denotes the average deviation from the the direct path.
  
  -> AUC denotes the area under the curve; meaning the geometric area between the actual trajectory and the direct path where the       area below the the direct path has been subtracted
  
  -> xpos_flips denotes the average number of directional changes along the x-axis
  
  -> RT denotes the average response time for the trials


# Visualisation

```{r}
# The function is used to aggregate mouse movement data from multiple trials or subjects into a single dataframe.
mt_aggregate(
  mt, use = "measures",
  use_variables = c("MAD", "AD", "AUC", "xpos_flips", "RT"),
  use2_variables = "Trial_Type",
  subject_id = "subject_nr"
)


```
```{r}
#plotting all trial curves with colors based on the condition
mt_plot(
  mt, use = "tn_trajectories",
  x = "xpos", y = "ypos", color = "Trial_Type") +
  theme_minimal() + labs(title = "Aligned time-normalized mouse-tracking data", 
                           x = "Position (x)",
                           y = "Position (y)")
```

```{r}
  #plotting the aggregated curves for the two conditions
mt_plot_aggregate(
  mt, use = "tn_trajectories",
  x = "xpos", y = "ypos", color = "Trial_Type",
  subject_id = "subject_nr") +
  theme_minimal() + labs(title = "Aligned aggregated time-normalized mouse trajectories", 
                           x = "Position (x)",
                           y = "Position (y)")
```
```{r}
mt_for_bind <- mt$measures %>% 
    select(-initiation_time, -mt_id) # remove duplicate data_collumns
visualisation <- cbind(mt_for_bind, mt$data) #bind data into one dataframe
visualisation_test_count <- visualisation %>% mutate(count = count_mouse+1)
rm(mt_for_bind) #remove extra intermediate dataframe

# create plot
RT_boxplot <- visualisation %>% 
  ggplot(aes(y=RT, x=Trial_Type, fill=Trial_Type)) +
    geom_boxplot()+
    geom_violin(alpha = .5)
 

# compute lower and upper whiskers
ylim1 = boxplot.stats(visualisation$RT)$stats[c(1, 5)]

# scale y limits based on ylim1
RT_boxplot + coord_cartesian(ylim = ylim1*1.05)


# ## Investigating columns with the same name
# mt$measures$mt_id == mt$data$mt_id
# mt$measures$initiation_time == mt$data$initiation_time                           
# 
# cbind(mt$measures$initiation_time, mt$data$initiation_time)
```


```{r}
# create plot
Correct_boxplot <- visualisation %>% 
  ggplot(aes(y=correct, x=Trial_Type, fill=Trial_Type)) +
    geom_histogram()+
    geom_violin(alpha = .5)
 

# compute lower and upper whiskers


# scale y limits based on ylim1
Correct_boxplot + coord_cartesian(ylim = ylim1*1.05)
```
```{r}
mean_RT_df <- visualisation %>%
  group_by(Trial_Type) %>%
  summarize(mean=mean(RT),SD=sd(RT))

visualisation %>% 
    ggplot(aes(x=RT, fill=Trial_Type)) +
    geom_density(alpha = .5)+
    xlim(300,1500)+
    geom_vline(data = mean_RT_df, aes(xintercept = mean, 
                                       color = Trial_Type), size=1, alpha = .3)+
    geom_text(data = mean_RT_df, aes(x=mean+ 400, 
                                     label=paste0("Mean:\n",round(mean,2)), y=0.0015))+
        geom_text(data = mean_RT_df, aes(x=mean + 380, 
                                     label=paste0("SD:\n",round(SD,2)), y=0.0013))+
    facet_wrap(~ Trial_Type)+
    theme_bw()+
    labs(title = "Density plots for reactiontime in ms", 
                           x = "Reactiontime in ms",
                           y = "Density")

```

```{r fitted line}
wrong_df <- df %>% 
    select(correct == 0)



```



```{r - modelling}
mGaus <- lmer(RT ~ Trial_Type + (1 | subject_nr)+ (1 | subject_nr)
              , data=visualisation)


mLog <- glmer(RT ~ Trial_Type + (1 | subject_nr)
              , family=gaussian( link="log")
              , data=visualisation)






summary(mLog)
anova(mGaus, mLog)
?isSingular


```

```{r}
wrong_answers_plot <- df_wrong %>% 
    ggplot(aes(x = count_mouse, fill = Trial_Type)) +
    geom_density(alpha = .5, trim=TRUE)+
    theme_bw()+ labs(title = "Density plot of wrong answers for each condition", 
                           x = "Trial",
                           y = "Density")


wrong_answers_plot

```



```{r}

df$correct <- as.logical(df$correct)



binomial <- glmer(correct ~ count_mouse*Trial_Type + (1 | subject_nr),
              data=df,
              family = "binomial")


summary(binomial)

```




